# netCDF files in R: Raster, Spatial objects

The aim of this tutorial is to provide a worked example (i.e., a function) of how to transform a regridded **netCDF** into a Raster object using R.

## Data import

Load the required packages.

```{r, message = FALSE, warning = FALSE, out.width = "40%", eval = FALSE}

install.packages(c("terra", "ncdf4", "ncdf4.helpers", "PCICt", 
                   "dplyr", "magrittr"))

# load packages
library(terra)
library(ncdf4)
library(ncdf4.helpers)
library(PCICt)
library(dplyr)
library(magrittr)

# # List of pacakges that we will use
#   list.of.packages <- c("raster", "data.table", "dplyr", "foreach", "doParallel", "ggplot2", "magrittr", "sf", "tidyselect")
# # If is not installed, install the pacakge
#   new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
#       if(length(new.packages)) install.packages(new.packages)
# # Load packages
#   lapply(list.of.packages, require, character.only = TRUE)

```

## Function to transform netCDF files into Raster objects

You can read netCDF using `raster::stack` or the `terra::rast` functions from the `raster` and `terra` packages. However, this code allows more control over the outputs.

```{r, eval = FALSE, out.width = "40%", eval = FALSE}
# NO GUARANTEES THAT CODE IS CORRECT
# Caveat Emptor! (Latin for "Let the buyer beware")
  
# Define generalities 
  nc = "data/BOAonMUR_SWIO_Y2003-M1-D1.nc"
  v = "temp_gradient" 
  x = "lon"
  y = "lat"

  nc <- ncdf4::nc_open(nc)
  dat <- ncdf4::ncvar_get(nc, v) # x, y, year
  dat[] <- dat
  rlon <- ncdf4::ncvar_get(nc, varid = x) %>% 
    range()
  rlat <- ncdf4::ncvar_get(nc, varid = y) %>% 
    range()
  X <- dim(dat)[1] # nolint
  Y <- dim(dat)[2] # nolint
  tt <- ncdf4.helpers::nc.get.time.series(nc,
                                          v = "time",
                                          time.dim.name = "time")
  tt <- as.POSIXct(tt)
  tt <- as.Date(tt)
  ncdf4::nc_close(nc)
# Make a raster with the right dims to fill with lat&lon
  rs <- terra::rast(nrow = Y, ncol = X, extent = terra::ext(c(rlon, rlat)))
  # Fix orientation of original data
  # [and then create a raster with this fix orientation]
    drs <- terra::xyFromCell(rs, 1:terra::ncell(rs)) %>%
      as_tibble()
  # Create raster stacks of depths for every month
    rs_list <- list() # to allocate results # nolint
    st <- terra::rast()
    for (i in 1:length(tt)) { # nolint
      dt1 <- dplyr::bind_cols(drs,
                              as.vector(dat[])) %>% # actually no need to loop but if you have more than just one year would be dat[, , i] , which is == X, Y, TIME
        magrittr::set_colnames(c("x", "y", v))
      dt1 <- terra::rast(dt1, type = "xyz")
      names(dt1) <- tt[i]
      st <- c(st, terra::flip(dt1))
      print(paste0(tt[i], " of ", length(tt)))
    }
    
```

Crop/Manipulate and project

```{r, eval = FALSE, out.width = "40%", eval = FALSE}
# The Mozambique Channel
  st <- terra::crop(st, terra::ext(c(30, 75, -35, -7)))
  terra::crs(st) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
```
